---
title: "x3p_processing"
author: "Cole Patten"
date: "2024-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## import libraries
```{r}
library(cmcR)
library(x3ptools)
library(ggplot2)
library(magrittr)
library(dplyr)
```

## Set path for reading x3ps
```{r}
read_path <- "/Users/colepatten/Desktop/NBIDE/cc/"

#record names of x3p files
files <- list.files(path=read_path, pattern = "\\.x3p$")
```

## function to view x3p image
```{r}
plot_im <- function(x3p_im){
  x3p_image(x3p_im, size=dim(x3p_im$surface.matrix), zoom=0.75, useNULL=TRUE)
}
```

## 
```{r}
cur_file = files[1]
im <- x3ptools::x3p_read(paste0(read_path, cur_file))
x3p_image(im, size=dim(im$surface.matrix), zoom=0.5, useNULL=TRUE)
```


## load all files, downsample high resolution files, then save
```{r}
# path to write downsampled files
write_path = "/Users/colepatten/Desktop/NBIDE/x3p_downsampled/"

# load x3ps, downsample by factor of 4, save
for (i in 1:108){
  cur_file = files[i]
  im <- x3ptools::x3p_read(paste0(read_path, cur_file))
  im_down <- im %>%
    x3p_sample(m=2)
  x3p_write(im_down, file=paste0(write_path, cur_file))
}

# load x3ps, save
for (i in 109:144){
  cur_file = files[i]
  im <- x3ptools::x3p_read(paste0(read_path, cur_file))
  x3p_write(im, file=paste0(write_path, cur_file))
}
```

## level all images to breech face place
```{r}
# new paths
read_path = "/Users/colepatten/Desktop/NBIDE/x3p_downsampled/"
write_path = "/Users/colepatten/Desktop/NBIDE/x3p_leveled/"

# load all x3ps, downsample higher resolution images by factor of 4, save
for (i in 1:length(files)){
  cur_file = files[i]
  im <- x3ptools::x3p_read(paste0(read_path, cur_file))
  im_level <- im %>%
    preProcess_ransacLevel()
  
  #x3p_image(im_level, size=dim(im$surface.matrix), zoom=0.75, useNULL=TRUE)
  x3p_write(im_level, file=paste0(write_path, cur_file))
}
```

## crop images to breech face only
```{r}
# new paths
read_path = "/Users/colepatten/Desktop/NBIDE/x3p_leveled/"
write_path = "/Users/colepatten/Desktop/NBIDE/x3p_cropped/"

#load images, remove FP, crop interior and exterior, save images
for (i in 1:length(files)){
  cur_file = files[i]
  im <- x3ptools::x3p_read(paste0(read_path, cur_file))
  im_cropped <- im %>%
    preProcess_removeFPCircle() %>%
    preProcess_erode(region="exterior", morphRadius=10) %>%
    preProcess_erode(region="interior", morphRadius=20) 
    
  
  #x3p_image(im_cropped, size=dim(im_cropped$surface.matrix), zoom=0.75, useNULL=TRUE)
  x3p_write(im_cropped, file=paste0(write_path, cur_file))
}
```

```{r}
read_path <- "/Users/colepatten/Desktop/NBIDE/x3p_downsampled/"
cur_file <- files[133]
bad_im <- x3ptools::x3p_read(paste0(read_path, cur_file))
bad_im2 <- bad_im %>%
  preProcess_ransacLevel()
plot_im(bad_im2)
```

```{r}
write_path <- "/Users/colepatten/Desktop/NBIDE/x3p_leveled/"
x3p_write(bad_im2, file=paste0(write_path, cur_file))

```

## we may have gone too far in cropping some images
## they should be checked and the bad ones should be manually cropped

```{r}
leveled_path <- "/Users/colepatten/Desktop/NBIDE/x3p_leveled/"
cropped_path <- "/Users/colepatten/Desktop/NBIDE/x3p_cropped/"

cur_file <- files[8]
num_files <- length(list.files(path=write_path, pattern = "\\.x3p$"))

im_leveled <- x3ptools::x3p_read(paste0(leveled_path, cur_file))
im_cropped <- x3ptools::x3p_read(paste0(cropped_path, cur_file))
im_pure <- im_cropped

print(paste(num_files, ":", cur_file))
x3pListPlot(list(im_leveled, im_cropped))
```

## if the automated cropping failed, manually edit the scan
```{r}
im_pure <- im_leveled %>%
  preProcess_erode(region="exterior", morphRadius=10) %>%
  #preProcess_erode(region="interior", morphRadius=30) %>%
  preProcess_erode(region="interior", morphRadius=20)

x3pListPlot(list(im_leveled, im_cropped, im_pure))
```

## save either im_leveled, im_cropped, or im_pure to x3p_pure
```{r}
write_path <- "/Users/colepatten/Desktop/NBIDE/x3p_pure/"
x3p_write(im_pure, file=paste0(write_path, cur_file))
```

## check all images against their pre-cropped version to see if there were errors made
```{r}
pure_path <- "/Users/colepatten/Desktop/NBIDE/x3p_pure/"
i <- 144
im_1 <- x3ptools::x3p_read(paste0(leveled_path, files[i]))
im_2 <- x3ptools::x3p_read(paste0(pure_path, files[i]))

x3pListPlot(list(im_1, im_2))
```

## level all images and apply Gaussian filter
```{r}
pure_path <- "/Users/colepatten/Desktop/NBIDE/x3p_pure/"
processed_path <- "/Users/colepatten/Desktop/NBIDE/x3p_processed/"

#load images, level, apply Gaussain filter, save
for (i in 1:length(files)){
  cur_file = files[i]
  im <- x3ptools::x3p_read(paste0(pure_path, cur_file))
  im_cropped <- im %>%
    preProcess_removeTrend(statistic = "quantile", tau = .5, method = "fn") %>%
    preProcess_gaussFilter()
  
  #x3p_image(im_cropped, size=dim(im_cropped$surface.matrix), zoom=0.75, useNULL=TRUE)
  x3p_write(im_cropped, file=paste0(processed_path, cur_file))
}
```

## check around to make sure data is consistent
```{r}
processed_path <- "/Users/colepatten/Desktop/NBIDE/x3p_processed/"

x3ps <- list()
for (i in 1:length(files)){
  cur_file = files[i]
  im <- x3ptools::x3p_read(paste0(processed_path, cur_file))
  x3ps <- append(x3ps, list(im))
  names(x3ps)[i] <- cur_file
}
```

## grab the x, y dimension of each scan, as well as the resolution.
## Check the minimum and maximum of each
```{r}
sizeXs <- integer(length=length(x3ps))
sizeYs <- integer(length=length(x3ps))
incrementXs <- integer(length=length(x3ps))
incrementYs <- integer(length=length(x3ps))

for (i in 1:length(x3ps)){
  sizeXs[i] <- x3ps[[i]]$header.info$sizeX
  sizeYs[i] <- x3ps[[i]]$header.info$sizeY
  incrementXs[i] <- x3ps[[i]]$header.info$incrementX
  incrementYs[i] <- x3ps[[i]]$header.info$incrementY

}
cat(paste(paste("sizeX: min:", min(sizeXs), "| max:", max(sizeXs)),
            paste("sizeY: min:", min(sizeYs), "| max:", max(sizeYs)),
            paste("incrementX: min:", min(incrementXs), "| max:", max(incrementXs)),
            paste("incrementY: min:", min(incrementYs), "| max:", max(incrementYs)),
            sep="\n"))
```
## lastly, we want to combine the images with data about which gun fired them ...
## in order to test our verification methods.
## here we will combine the data, then export as a csv to use in python
```{r}
library(readxl)
```

```{r}
excel_path <- "/Users/colepatten/Desktop/NBIDE/StudyInfo.xlsx"
sheets <- excel_sheets(excel_path)
print(sheets)
```

```{r}
read_and_clean_sheet <- function(file_path, sheet_name) {
  # Read the sheet, starting from the second row
  raw_data <- read_excel(file_path, sheet = sheet_name, skip = 0)
  
  # Set the column names to the first row of raw_data and remove that row
  colnames(raw_data) <- raw_data[1, ]
  clean_data <- raw_data[-1, ]
  
  # Return the cleaned data frame
  return(clean_data)
}
```

```{r}
# Read and clean each sheet into a list of data frames
data_list <- lapply(sheets, function(sheet) {
  read_and_clean_sheet(excel_path, sheet)
})

# Name the list elements with the sheet names
names(data_list) <- sheets

# Print the first few rows of each data frame to check the import
lapply(data_list, head)
```

## rename columns in the orignal data frame in order to allow for merging
```{r}
names(data_list$Firearm)[names(data_list$Firearm) == "Firearm Name"] <- "Firearm"
names(data_list$Firearm)[names(data_list$Firearm) == "Brand"] <- "Firearm_Brand"
names(data_list$Firearm)[names(data_list$Firearm) == "Model"] <- "Firearm_Model"

names(data_list$`Cartridge Case`)[names(data_list$`Cartridge Case`) == "Specimen ID"] <- "Specimen"
names(data_list$`Cartridge Case`)[names(data_list$`Cartridge Case`) == "Firearm Name"] <- "Firearm"
names(data_list$`Cartridge Case`)[names(data_list$`Cartridge Case`) == "Brand"] <- "Cartridge_Brand"

names(data_list$`Cartridge Case Measurement`)[names(data_list$`Cartridge Case Measurement`) == 
                                                "Specimen Name"] <-
  "Specimen"
names(data_list$`Cartridge Case Measurement`)[names(data_list$`Cartridge Case Measurement`) == 
                                                "File Name"] <-
  "File"

```

## reduce data frames to only the need columns
```{r}
firearm_df <- data_list$Firearm[1:12, c("Firearm", "Firearm_Brand", "Firearm_Model")]
cartridge_df <- data_list$`Cartridge Case`[1:144,c("Specimen", "Firearm", "Cartridge_Brand")]
file_df <- data_list$`Cartridge Case Measurement`[1:144, c("Specimen", "File")]
```

## merge into new data frame with relevant data
```{r}
dat <- right_join(x=firearm_df,
           y=cartridge_df,
           by = "Firearm",
           relationship="one-to-many")
dat <- inner_join(x=dat,
           y=file_df,
           by="Specimen",
           relationship="one-to-one")
```

## save this file
```{r}
write_path <- "/Users/colepatten/Desktop/NBIDE/info.csv"
write.csv(dat, write_path, row.names = FALSE)
```

## now we want to convert the x3ps to data frames to send to python
```{r}
write_path <- "/Users/colepatten/Desktop/NBIDE/x3p_dataframe/"
for (i in 1:length(x3ps)){
  specimen_name <- dat[dat$File == names(x3ps)[i], ]$Specimen
  write.csv(x3p_to_df(x3ps[[i]]),
            paste0(write_path, specimen_name, ".csv"),
            row.names = FALSE)
}

```
```{r}
paste0(write_path, specimen_name, ".csv")
```

